---
title: "pefloxacin typhi analysis and figures"
author: "Philip Ashton"
date: "`r Sys.Date()`"
output: html_document
---

# Libraries

```{r message=FALSE}
library(tidyverse)
library(patchwork)
library(kableExtra)
library(broom)
library(sjPlot)

source('pefloxacin_typhi_functions.R')
```

# Read in data

```{r}
cases_bc_qrdr_prescrip_microconf_handle <- 'data/2022.09.14.pef_typhi.cases_bc_qrdr_prescrip_microconf.csv'
qech_pef_by_month_handle <- 'data/2022.09.14.qech_pef_results_per_month.csv' 
strataa_tyvac_qrdr_typhi_handle <- 'data/2022.09.15.strataa_tyvac_qrdr_typhi.one_per_patient.csv'
tree_summaries_handle <- 'data/2022.09.16.tree_summaries.csv'
sequencing_confirmed_typhi_opp_handle <- 'data/2022.09.14.sequencing_confirmed_typhi_opp.csv'
```

Data has been pulled in from various sources and cleaned using `pefloxacin_typhi_data_cleaning.Rmd`. The clean data has been saved as `data/cases_bc_qrdr_prescrip_microconf_handle`.

```{r}
sequencing_confirmed_typhi_opp <- read_csv(sequencing_confirmed_typhi_opp_handle)
entire_recruitment_period <- read_csv(cases_bc_qrdr_prescrip_microconf_handle)
```

# STRATAA/TyVAC data summary

The number of sequencing confirmed typhoid cases and number of qrdr (sequencing confirmed), number of prescriptions of cipro, number of blood cultures, and number of microbiologically confirmed typhoid.

```{r}
typhoid_n <- sum(entire_recruitment_period$total_seq_typhoid_n)
typhi_n <- sum(entire_recruitment_period$total_seq_typhi_n)
prescrip_n <- sum(entire_recruitment_period$total_prescrip_n)
qrdr_n <- sum(entire_recruitment_period$total_qrdr_n)
bc_n <- sum(entire_recruitment_period$total_bc_n)
micro_conf_opp_n <- sum(entire_recruitment_period$micro_confirmed_opp_n)
micro_conf_all_n <- sum(entire_recruitment_period$micro_confirmed_all_n)
```


In total, over the entire study period there were:

* `r micro_conf_opp_n` cases of microbiologically confirmed typhoid (one per patient).
* `r micro_conf_all_n` isolates S. Typhi.
* `r typhoid_n` cases of sequencing confirmed typhoid 
* `r typhi_n` cases of sequencing confirmed typhi (multiple isolates per patient in some cases)
* `r prescrip_n` prescriptions of cipro
* `r qrdr_n` QRDR typhoid (i.e. resistant)
* `r format(bc_n, scientific=FALSE)` blood cultures


# Investigating associations

After discussion with Marc, want to do one model with all the data from the entire study period. It doesn't matter about the difference in recruitment criteria between strataa and tyvac because the important thing is the number of qrdr typhi per typhoid case, and the method of detecting those was consistent.

```{r glm model}

m1 <- glm(cbind(total_qrdr_n, total_seq_typhoid_n-total_qrdr_n) ~ lag(total_prescrip_n),  
         data = entire_recruitment_period, 
         family = "binomial")

#summary(m1)
#m1_summary <- broom.mixed::tidy(m1, exponentiate = TRUE, conf.int = TRUE)
#m1_summary  %>% kbl() %>% kable_styling()

tab_model(m1, dv.labels = c('Proportion of S. Typhi that have a QRDR mutant ~ Number of ciprofloxacin prescriptions'))

```

Checking for association between proportion of QRDR mutants and teh total number of blood cultures.

```{r}
m1 <- glm(cbind(total_qrdr_n, total_seq_typhoid_n-total_qrdr_n) ~ total_bc_n,  
         data = entire_recruitment_period, 
         family = "binomial")

summary(m1)
```

# Graphs
## Figure 1

This is a figure of the number of S. Typhi at QECH that were cipro sensitive and DCS.

```{r Figure1, fig.width = 8, fig.height = 4}

qech_pef_by_month <- read_csv(qech_pef_by_month_handle)

qech_pef_by_month_for_plotting <- qech_pef_by_month
df <- qech_pef_by_month_for_plotting %>% group_by(Month) %>% summarise(n = sum(n)) %>% mutate(n = n / 50)

qech_n_graph <- ggplot(qech_pef_by_month_for_plotting, aes(x = Month, y = n, fill = `Cipro MIC res?`)) + 
  geom_bar(stat = "identity") + 
  ylab("Count") +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=14)) +
  ggtitle('A)') + 
  scale_fill_discrete(name = "Ciprofloxacin\nMIC result",  labels = c("Decreased\nCiprofloxacin\nSusceptibility", "Resistant", "Sensitive"))

qech_prop_graph <- ggplot() + 
  geom_bar(qech_pef_by_month_for_plotting, mapping = aes(x = Month, y = n, fill = `Cipro MIC res?`), position = 'fill', stat = "identity") + 
  geom_line(df, mapping = aes(x = Month, y = n)) +
  # secondary axis plot https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
  scale_y_continuous("Proportion", breaks = (seq(0,1, by = 0.2)), sec.axis = sec_axis(~ . * 50, name = "# S. Typhi isolates", breaks = seq(0, 50, by = 10))) +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=14)) +
  scale_fill_discrete(name = "Ciprofloxacin\nMIC result", labels = c("Decreased\nCiprofloxacin\nSusceptibility", "Resistant", "Sensitive"))

qech_prop_graph

#qech_n_graph / qech_prop_graph #+ plot_layout(heights = c(100, 100))

```

Test for trend in proportions

```{r}
qech_pef_by_month_w <- pivot_wider(qech_pef_by_month, names_from = `Cipro MIC res?`, values_from = n) %>% replace_na(list(DCS = 0, R= 0)) %>% mutate(n = S + DCS + R, e = DCS + R)

prop.trend.test(qech_pef_by_month_w$e, qech_pef_by_month_w$n)
```

### Numbers for Figure 1

```{r}
total_qech_typhi_tested <- sum(qech_pef_by_month$n)
total_qech_typhi_dcs_or_r <- filter(qech_pef_by_month, `Cipro MIC res?` == 'DCS' | `Cipro MIC res?` == 'R' ) %>% pull(n) %>% sum
```

There were `r total_qech_typhi_tested` total qech typhi tested, and `r total_qech_typhi_dcs_or_r` DCS or R.


## Figure 2

Read in the info about the qrdr samples.

```{r}
strataa_tyvac_qrdr_typhi <- read_csv(strataa_tyvac_qrdr_typhi_handle, col_types = cols(`Properly formatted date` = col_date(format = "%Y-%m-%d")))
strataa_tyvac_qrdr_typhi$Month <- as.Date(cut(strataa_tyvac_qrdr_typhi$`Properly formatted date`, breaks = "month"))
```

Graphs of the counts of typhi cases, qrdr cases, cipro prescriptions and blood cultures from the entire strataa/tyvac recruitment period.


```{r, fig.width = 6, fig.height = 12}
entire_period_tpqb_graph <- plot_typh_presc_qrdr_bc_n(entire_recruitment_period, strataa_tyvac_qrdr_typhi) # tpqb stands for typhoid, prescriptions, qrdr and blood cultures
entire_period_tpqb_graph
```

### Numbers for Figure 2

```{r}
qrdr_before_sept019 <- strataa_tyvac_qrdr_typhi %>% filter(`Properly formatted date` < '2019-09-01')
qrdr_sept_oct2019 <- strataa_tyvac_qrdr_typhi %>% filter(`Properly formatted date` >= '2019-09-01' & `Properly formatted date` <= '2019-10-31')

mutations_before_sept2019 <- qrdr_before_sept019 %>% group_by(QRDR) %>% summarise(n = n())
mutations_sept_oct2019 <- qrdr_sept_oct2019 %>% group_by(QRDR) %>% summarise(n = n())

#before_sept19 <- entire_recruitment_period %>% filter(`Month` < '2019-09-01')
#after_sept19 <- entire_recruitment_period %>% filter(`Month` >= '2019-09-01')

before_after_sept19 <- entire_recruitment_period %>% mutate(before_after_sept19 = ifelse(Month < '2019-09-01', 'before_sept19', 'after_sept19')) %>% group_by(before_after_sept19, ) %>% summarise(total_seq_typhoid_n = sum(total_seq_typhoid_n), total_qrdr_n = sum(total_qrdr_n)) %>% mutate(pct_qrdr = (total_qrdr_n / total_seq_typhoid_n) * 100)
before_after_sept19 %>% kbl() %>% kable_styling()



```

Get the Number of typhoid cases before/after Dec 2018

```{r}

entire_recruitment_period %>% mutate(bef_aft_dec18 = ifelse(Month <= "2018-12-01", 'before_dec18', 'after_dec18')) %>% group_by(bef_aft_dec18) %>% summarise(min_tyhpoid = min(total_seq_typhoid_n), max_tyhpoid = max(total_seq_typhoid_n), median_tyhpoid = median(total_seq_typhoid_n)) %>% kbl() %>% kable_styling()

```


Get the max, min and average number of prescriptions and qrdr cases before/after May 2019.

```{r}
entire_recruitment_period %>% mutate(bef_aft_may19 = ifelse(Month <= "2019-05-01", 'before_may19', 'after_may19')) %>% group_by(bef_aft_may19) %>% summarise(min_prescrip = min(total_prescrip_n), max_prescrip = max(total_prescrip_n), median_prescrip = median(total_prescrip_n), min_qrdr = min(total_qrdr_n), max_qrdr = max(total_qrdr_n), mean_qrdr = mean(total_qrdr_n)) %>% kbl() %>% kable_styling()

```


```{r}
x <- c(4, 6)
y <- c(171, 12)
prop.test(x, y, p = NULL, alternative = "two.sided")
```


There were `r nrow(qrdr_before_sept019)` QRDR Typhi isolated between 1st October 2016 and 31th August 2019.

There were `r nrow(qrdr_sept_oct2019)` QRDR Typhi isolated between 1st September 2019 and 31st October 2019.

Before Sept2019
`r mutations_before_sept2019  %>% kbl() %>% kable_styling()`
Sept-Oct 2019
`r mutations_sept_oct2019  %>% kbl() %>% kable_styling()`


# Pheno-geno comparison


```{r}
pheno_or_geno <- sequencing_confirmed_typhi_opp %>% filter((`Pefloxacin 5` == 'R' | `Pefloxacin 5` == 'I') | (`Cipro MIC res?` == 'DCS' | `Cipro MIC res?` == 'R') | !is.na(QRDR)) %>% select(c(`Request Number`, `Pefloxacin 5`, `Ciprofloxacin MIC`, `Cipro MIC res?`, QRDR, any_qrdr)) %>% replace_na(list(QRDR = 'None', any_qrdr = 'None'))

table(pheno_or_geno$any_qrdr, pheno_or_geno$`Cipro MIC res?`)

```


# Summaries of the numbers in each phylo

```{r}
all_tree_summaries <- read_csv(tree_summaries_handle)
context <- all_tree_summaries %>% filter(tree == 'malawian_and_context') %>% filter(Country != 'Malawi')

number_of_nonmalawian_context_genomes <- sum(context$n)
```

There were `r number_of_nonmalawian_context_genomes` non-Malawian contextual genomes included in the analysis.

Table of the number of genomes included from each location in each tree.

`r all_tree_summaries %>% kbl() %>% kable_styling()`


# Supplmentary Figure 1 - Plot number of sequenced and unsequenced typhi cases

```{r}
seq_unseq_cases <- entire_recruitment_period %>% select(Month, total_seq_typhoid_n, micro_confirmed_opp_n) %>% mutate(unsequenced = micro_confirmed_opp_n - total_seq_typhoid_n)
seq_unseq_cases_l <- seq_unseq_cases %>% select(Month, total_seq_typhoid_n, unsequenced) %>% pivot_longer(!Month, names_to = "type", values_to = "Number")
ggplot(seq_unseq_cases_l, aes(x = Month, y = Number, fill = type)) +
  geom_bar(stat = "identity")
```






